Improving disaggregation models of malaria incidence by ensembling non-linear models of prevalence

### 2019-10-25 tim-lucas-model-ensembles.tex
New figures and things.

\begin{table}[t!]
\caption{Pearson correlations between observed and predicted values. }
\centering
\begin{tabular}{llrrrr}
Country &  Cross-validation & Enviro &  ML\textsubscript{l} &  Enviro + ML\textsubscript{l} \\
\hline 
 Colombia & Random &  0.59 & 0.59 & \textbf{0.61} \\
 Colombia &  Spatial &  0.12 &  \textbf{0.33} &  0.25\\
 Madagascar & Random &   \textbf{0.70} &  0.68 & 0.69 \\
 Madagascar &  Spatial &  0.22 &  \textbf{0.55} & 0.18\\
 Senegal &  Random &  \textbf{0.58} &  0.51 & 0.57 \\
 Senegal &  Spatial &  \textbf{0.63} &  0.51 &  \textbf{0.63} \\
 Indonesia &  Random &  0.58 &  0.48 &  \textbf{0.59} \\
 Indonesia &  Spatial &  \textbf{0.63} &  0.44 &  0.51 \\
\end{tabular}
\label{t:results}
\end{table}

@@

\begin{table}[t!]
\caption{Pearson correlations between observed and predicted values. }
\centering
\begin{tabular}{llrrrr}
Country &  Cross-validation & Enviro &   ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
\hline 
 Colombia & Random &  \textbf{0.59} &0.55 & 0.58 \\
 Colombia &  Spatial &  0.12 &  \textbf{0.33} & 0.25\\
 Madagascar &  Random &  \textbf{0.70} &  0.67 & 0.68 \\
 Madagascar &  Spatial &  0.22 &  \textbf{0.55} & 0.18\\
 Senegal &  Random &  \textbf{0.58} &  0.50 & 0.49 \\
 Senegal &  Spatial &  \textbf{0.63} & 0.51 & \textbf{0.63} \\
 Indonesia &  Random &  \textbf{0.52} & 0.32 & 0.46 \\
 Indonesia &  Spatial &  \textbf{0.63} & 0.44 & 0.51 \\
\end{tabular}
\label{t:results2}
\end{table}
Figures~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for all countries and all models.
Figures~\ref{f:scatter2} shows scatter plots of the model performance under spatial cross-validation for all countries and all models.
It can be seen that without environmental covariates, the models in Madagascar fail to predict very high or very low values correctly.
Figure~\ref{f:map} shows the input data and out-of-sample predictions of the ML\textsubscript{l} model in Colombia.
Table~\ref{t:coverage} shows the out-of-sample coverage of the 80\% credible intervals for all models, countries and cross-validation schemes.
The coverage in Colombia is very poor with no models achieving a coverage above 0.4.
Otherwise the coverage is generally acceptable with most values lying between 0.7 and 0.9.
Over there is a general tendancy for models to be slightly over confident.

@@

\begin{figure}
\centering
\includegraphics[trim={0 30mm 0 40mm}, width = 0.8\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
\caption{
  Left: Observed data for Colombia (grey for zero incidence). Right: Out-of-sample predictions for the random cross-validation, machine learning only model. For each cross-validation fold, predictions are made for the held out data which are then combined to make a single surface.
}
\label{f:map}
\end{figure}

\begin{table}[h!]
\caption{Covarage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
\centering
\begin{tabular}{llrrrrr}
Country &  CV & Enviro & ML\textsubscript{l} &  Enviro + ML\textsubscript{l} & ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
\hline 
 Colombia & Random & \textbf{0.28} & \textbf{0.28} & \textbf{0.29} & \textbf{0.28} & \textbf{0.30} \\
 Madagascar &  Random & 0.80 & 0.77& 0.75& 0.77& 0.76 \\
 Senegal & Random &0.79 & 0.79& 0.79& 0.79& 0.82 \\
 Indonesia & Random &0.80 & 0.81& 0.78& 0.79& 0.77  \\
 Colombia &  Spatial & \textbf{0.30} & \textbf{0.33}  & \textbf{0.33} & \textbf{0.33} & \textbf{0.34}  \\
 Madagascar & Spatial & \textbf{0.65} & 0.74& 0.70& 0.75 & 0.76   \\
 Senegal & Spatial & 0.85 & \textbf{0.91}& 0.85& \textbf{0.94} & 0.85  \\
  Indonesia & Spatial & 0.80 & 0.78& 0.78& 0.76& 0.75  \\
\end{tabular}
\label{t:coverage}
\end{table}

\begin{table}
\caption{Machine learning model results and fitted parameters (i.e.\thinspace model weights) of the machine learning predictions only models (local only and global only). }
\centering
\small
\begin{tabular}{ll|rr|rr}
     Country          & Model &      RMSE\textsubscript{l} & $\beta_l$ & RMSE\textsubscript{g} & $\beta_g$ \\ \hline
Madagascar & RF & \textbf{0.100} & 0.538 &  \textbf{0.169} & \textbf{0.529}\\
Madagascar & GBM & 0.105 & \textbf{0.570} & 0.178& 0.432 \\
Madagascar & enet & 0.116 & 0.301 &0.233 & 0.262 \\
Madagascar & nnet & 0.113 & 0.033 &0.220 & 0.364 \\
Madagascar & ppr & 0.109 & 0.469 & 0.205 &  0.403\vspace{0.3cm}\\ 
Colombia & RF & \textbf{0.068} & 0.625 &  \textbf{0.169} & 0.180\\
Colombia & GBM & 0.073 & \textbf{0.952} & 0.178& -0.218  \\
Colombia & enet & 0.070 & 0.219 &0.233 & 0.183 \\
Colombia & nnet & 0.070 & 0.129 &0.220 & 0.527 \\
Colombia & ppr & 0.070 & 0.0.667 & 0.205 &  \textbf{0.546}\vspace{0.3cm}\\
Senegal & RF & \textbf{0.092} & \textbf{0.339} & \textbf{0.169} & \textbf{0.425} \\
Senegal & GBM & 0.099 & 0.261& 0.178& 0.408 \\
Senegal & enet& 0.103 & 0.344  &0.233 & 0.205 \\
Senegal & nnet & 0.099 & 0.254 &0.220 & 0.190 \\
Senegal & ppr & 0.098 & 0.268& 0.205 &  0.126\vspace{0.3cm}\\
Indonesia & RF& \textbf{0.081} & 0.447 & \textbf{0.169} & 0.178\\
Indonesia & GBM & 0.085 & 0.357 & 0.178& 0.289 \\
Indonesia & enet & 0.091 & 0.303 &0.233 & \textbf{0.526} \\
Indonesia & nnet & 0.089 & \textbf{0.506} &0.220 & 0.316 \\
Indonesia & ppr & 0.089 & 0.364 & 0.205 &  0.089\vspace{0.3cm}\\

>>>>

\begin{table}[t!]
\caption{Pearson correlations between observed and predicted values. }
\centering
\begin{tabular}{llrrrr}
Country &  Cross-validation & Enviro &  ML\textsubscript{l} &  Enviro + ML\textsubscript{l} \\
\hline 
 Colombia & Random &  0.59 & 0.59 & \textbf{0.61} \\
 Colombia &  Spatial &  0.12 &  \textbf{0.33} &  0.25\\
 Madagascar & Random &   \textbf{0.70} &  0.68 & 0.69 \\
 Madagascar &  Spatial &  0.22 &  \textbf{0.55} & 0.18\\
 Senegal &  Random &  \textbf{0.58} &  0.51 & 0.57 \\
 Senegal &  Spatial &  0.63 &  0.51 &  0.63 \\
 Indonesia &  Random &  0.52 &  0.48 &  \textbf{0.59} \\
 Indonesia &  Spatial &  0.45 &  0.44 &  \textbf{0.51} \\
\end{tabular}
\label{t:results}
\end{table}

@@

\begin{table}[t!]
\caption{Pearson correlations between observed and predicted values. }
\centering
\begin{tabular}{llrrrr}
Country &  Cross-validation & Enviro &   ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
\hline 
 Colombia & Random &  \textbf{0.59} &0.55 & 0.58 \\
 Colombia &  Spatial &  0.12 &  0.12 & \textbf{0.33}\\
 Madagascar &  Random &  \textbf{0.70} &  0.67 & 0.68 \\
 Madagascar &  Spatial &  0.22 &  0.51 & \textbf{0.55}\\
 Senegal &  Random &  \textbf{0.58} &  0.50 & 0.49 \\
 Senegal &  Spatial &   \textbf{0.63} & 0.55 & 0.52 \\
 Indonesia &  Random &  \textbf{0.52} & 0.32 & 0.46 \\
 Indonesia &  Spatial &  0.45 & 0.41 & 0.45 \\
\end{tabular}
\label{t:results2}
\end{table}
Figures~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for all countries and all models.
Figures~\ref{f:scatter2} shows scatter plots of the model performance under spatial cross-validation for all countries and all models.
It can be seen that without environmental covariates, the models in Madagascar fail to predict very high or very low values correctly.
Figure~\ref{f:map} shows the input data and out-of-sample predictions of the ML\textsubscript{l} model in Colombia.
Table~\ref{t:coverage} shows the out-of-sample coverage of the 80\% credible intervals for all models, countries and cross-validation schemes.
The coverage in Colombia is very poor with no models achieving a coverage above 0.4.
Otherwise the coverage is generally acceptable with most values lying between 0.7 and 0.9.
Overall there is a general tendancy for models to be slightly over confident.

@@

\begin{figure}
\centering
\includegraphics[width = 1\textwidth]{figs/col_comparison_map.png} %\caption{Indonesia spatial crossvalidation} 
\caption{
  Left: Observed data for Colombia (grey for zero incidence). Middle: Out-of-sample predictions for the spatial cross-validation, environmental covariates only model. Right: Out-of-sample predictions for the spatial cross-validation, local machine learning only model. For each cross-validation fold, predictions are made for the held out data which are then combined to make a single surface.
}
\label{f:map}
\end{figure}

\begin{table}[t!]
\caption{Covarage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
\centering
\begin{tabular}{llrrrrr}
Country &  CV & Enviro & ML\textsubscript{l} &  Enviro + ML\textsubscript{l} & ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
\hline 
 Colombia & Random & \textbf{0.28} & \textbf{0.28} & \textbf{0.29} & \textbf{0.28} & \textbf{0.30} \\
 Madagascar &  Random & 0.80 & 0.77& 0.75& 0.77& 0.76 \\
 Senegal & Random &0.79 & 0.79& 0.79& 0.79& 0.82 \\
 Indonesia & Random &0.80 & 0.81& 0.78& 0.79& 0.77  \\
 Colombia &  Spatial & \textbf{0.30} & \textbf{0.33}  & \textbf{0.33} & \textbf{0.33} & \textbf{0.34}  \\
 Madagascar & Spatial & \textbf{0.65} & 0.74& 0.70& 0.75 & 0.76   \\
 Senegal & Spatial & 0.85 & \textbf{0.91}& 0.85& \textbf{0.94} & 0.85  \\
  Indonesia & Spatial & 0.80 & 0.78& 0.78& 0.76& 0.75  \\
\end{tabular}
\label{t:coverage}
\end{table}

\begin{table}[t!]
\caption{Machine learning model results and fitted parameters (i.e.\thinspace model weights) of the machine learning predictions only models (local only and global only). }
\centering
\small
\begin{tabular}{ll|rr|rr}
     Country          & Model &      RMSE\textsubscript{l} & $\beta_l$ & RMSE\textsubscript{g} & $\beta_g$ \\ \hline
Madagascar & RF & \textbf{0.100} & 0.538 &  \textbf{0.169} & \textbf{0.529}\\
Madagascar & GBM & 0.105 & \textbf{0.570} & 0.178& 0.432 \\
Madagascar & enet & 0.116 & 0.301 &0.233 & 0.262 \\
Madagascar & nnet & 0.113 & 0.033 &0.220 & 0.364 \\
Madagascar & ppr & 0.109 & 0.469 & 0.205 &  0.403\vspace{0.3cm}\\ 
Colombia & RF & \textbf{0.068} & 0.625 &  \textbf{0.169} & 0.180\\
Colombia & GBM & 0.073 & \textbf{0.952} & 0.178& -0.218  \\
Colombia & enet & 0.070 & 0.219 &0.233 & 0.183 \\
Colombia & nnet & 0.070 & 0.129 &0.220 & 0.527 \\
Colombia & ppr & 0.070 & 0.667 & 0.205 &  \textbf{0.546}\vspace{0.3cm}\\
Senegal & RF & \textbf{0.092} & \textbf{0.339} & \textbf{0.169} & \textbf{0.425} \\
Senegal & GBM & 0.099 & 0.261& 0.178& 0.408 \\
Senegal & enet& 0.103 & 0.344  &0.233 & 0.205 \\
Senegal & nnet & 0.099 & 0.254 &0.220 & 0.190 \\
Senegal & ppr & 0.098 & 0.268& 0.205 &  0.126\vspace{0.3cm}\\
Indonesia & RF& \textbf{0.081} & 0.447 & \textbf{0.169} & 0.178\\
Indonesia & GBM & 0.085 & 0.357 & 0.178& 0.289 \\
Indonesia & enet & 0.091 & 0.303 &0.233 & \textbf{0.526} \\
Indonesia & nnet & 0.089 & \textbf{0.506} &0.220 & 0.316 \\
Indonesia & ppr & 0.089 & 0.364 & 0.205 &  0.089\vspace{0.3cm}\\
